---
title: "Saving your output"
format: 
  revealjs:
    slide-number: true
    embed-resources: true
editor: source
---

## Saving output as pdf (1 of 4)

```{r}
#| label: 01-07-documentation

tx <- readLines("simon-5507-01-save-output.sas")
```

```{}
`r extract(tx, "^\\*", ";")`
```

::: notes

Let's look at a modified program that stores your output as a pdf file.

Here is the documentation header.

:::

## Saving output as pdf (2 of 4)

```{}
`r extract(tx, "^libname", "output.pdf")`
```

::: notes

The ods statement should be placed near the top of the code, certainly before any SAS procedure that produces output.

:::

## Saving output as pdf (3 of 4)

```{}
`r extract(tx, "^data", "^;$")`
```

::: notes

The ods statement should be placed near the top of the code, certainly before any SAS procedure that produces output.

:::

## Saving output as pdf (4 of 4)

```{r}
#| label: 01-07-close

tx %>% str_which("proc") -> i4
tx %>% str_which("close;") -> j4

tx[i4:j4] %>%
  paste0(collapse="\n") -> tx4  
```


```{}
`r extract(tx, "proc print", "close")`
```

::: notes

Near the bottom, you turn off the output. Place this AFTER the last SAS procedure that produces output.

:::

## Live demonstration (4 of 5)

