---
title: "Sweep html files from src to results"
format: 
  html:
    embed-resources: true
execute: 
  echo: false
---

## File details

This program moves files from the src folder to results folder in each of the modules. It was written by Steve Simon on 2025-01-08 and is placed in the public domain. 

## Remind on how to print pdf files with speaker notes

To print speaker notes, add

-   ?print-pdf&showNotes=separate-page 

to the end of the URL. Then open the print dialog box (ctrl-P) and choose

-   Save as PDF

for the printer. Other options, like Microsoft Print to PDF, do not seem to work.

```{r}
#| label: setup
#| message: false
#| warning: false

library(glue)
library(tidyverse)
```

## Identify html files

```{r}
#| label: identify

identify_files <- function(subdir, ext) {
  list.files(
    path=glue("{subdir}/src"),
    pattern=ext)
}
```

## Move html files

```{r}
#| label: sweep

sweep_files <- function(subdir, file_name) {
  old <- glue("{subdir}/src/{file_name}")
  new <- glue("{subdir}/results/{file_name}")
  file.rename(from=old, to=new)
}
```

## Loop across folders

```{r}
#| label: loop

z99 <- c(glue("0{1:9}"), as.character(10:99))

loop_folders <- function(n) {
  modules <- z99[1:n]
  for (s in modules) {
  	s00 <- glue("{s}/src")
    print(glue("Searching in module {s00}"))
    file_list <- character(0)
    for (ext in c("docx", "html", "pdf", "pptx")) {
      file_list <- c(file_list, identify_files(s, ext))
    }
    if (length(file_list) == 0) next
    print(glue("---{s}/src/{file_list}"))
    for (f in file_list) {
      sweep_files(s, f)
    }
  }
}
```

## 5507

```{r}
#| label: r-5507

loop_folders(5)
```
