---
title: "convert_to_markdown"
format: 
  html:
    embed-resources: true
editor: source
---

```{r}
#| label: setup
#| message: false
#| warning: false

library(tidyverse)

marker_heads <- "^\\* "
prefix_heads <- "\n\n## ",
postfix_heads <- "\n\n"

marker_notes <- "^\\* Comments on the code: "
prefix_notes <- "\n\n::: {.notes}\n\n"
postfix_notes <- 

prefix_sas <- "```{}\n"
postfix_sas <- "\n\n:::\n\n"

semicolon <- ";" 
```

```{r}
#| label: read

read_code <- function(code_file) {
  # reads sas file and splits into chunks
  read_lines(paste0(code_file, ".sas")) |>
  paste0(collapse="\n") |>
  str_split(pattern="\n\n\n") |>
  unlist()
}
```

```{r}
#| label: parse-heads

parse_heads <- function(chunks) {
  # find chunks which look like heads, add prefix and postfix
  heads <- str_detect(chunks, marker_heads) & !str_detect(chunks, marker_notes)

  chunks[heads] <- str_remove(chunks[heads], marker_heads)
  chunks[heads] <- str_remove(chunks[heads], semicolon)

  chunks[heads] <- paste0(
    prefix_heads,
    chunks[heads],
    postfix_heads)
  
  return(chunks)
}

parse_notes <- function(chunks) {
  # find chunks which look like notes, add prefix and postfix
  notes <- str_detect(chunks, marker_notes)

  chunks[notes] <- str_remove(chunks[notes], market_notes)
  chunks[notes] <- str_remove(chunks[notes], semicolon)
  
  chunks[notes] <- paste0(
    prefix_notes,
    chunks[notes],
    postfix_notes)
}

parse_sas <- function(chunks) {
  # find chunks which look like sas, add prefix and postfix
  sas <- !str_detect(marker_heads)

  chunks[sas] <- paste0(
    prefix_sas,
    chunks[sas],
    postfix_sas)

  return(chunks)
  
}

parse_code <- function(chunks) {


  sas <- !notes & !heads


}
```


```{r}
#| label: write

write_markdown <- function(chunks) {
  cat("\n\n## Documentation header\n\n")
  cat(chunks, sep="\n")
}
```
