---
title: "sas-05-01"
format: 
  revealjs:
    slide-number: true
    embed-resources: true
editor: source
---

```{r}
#| message: false
#| warning: false
#| label: 01-setup

# I will not use too much R in a class called
# "Introduction to SAS," but am including this
# chunk just in case I need something from R.

library(glue)
library(tidyverse)
```

{{< include ../../general/src/chunk_sas_function.qmd >}}

```{r}
#| label: 05-01-read

sas <- chunk_sas("simon-5507-05-demo.sas") 
```

## Overview

-   Mix of categorical and continuous variables
    -   Review methods for continuous variables
    -   Review methods for categorical variables
    -   New methods for mix of categorical and continuous variables

::: notes

Today, you will analyze some data sets that have a mix of categorical and continuous variables. The first data set looks at pulmonary function in a group of children.

You can find a description of this data set at

  http://jse.amstat.org/datasets/fev.txt

:::

## Labels and formats

-   Document your data
    -   Formats for number codes
    -   Labels for all variables
    
::: notes

Documenting your code is important. If your categorical data uses number codes, make sure to define what those codes represent using proc format.

Also use labels for all variables, with the possible exception of simple variables that are self-explanatory with the variable name alone.

:::

`r sas[1]`

`r sas[2]`

`r sas[3]`

`r sas[4]`

`r sas[5]`
