---
title: "sas-06-02"
format: 
  revealjs:
    slide-number: true
    embed-resources: true
editor: source
---

```{r}
#| label: 06-02-read
#| message: false
#| warning: false

library(tidyverse)

text_data <- read_lines("todd-5507-06-demo.sas")
```

## 06-02 slide 0

Testing slide text

::: {.notes}

Testing notes text

:::

## 06-02 slide 1

```{r}

cat(text_data[1:5], sep="\n")
>>>>>>> 97c7faf4f68264102f4cdf44f1276f341fe159a5
```

```{r}
#| label: display-chunks
#| asis: true

<<<<<<< HEAD
for (i in 1:length(code_chunks)) {
  cat("\n\n## Code chunk ")
  cat(i)
  cat("\n\n")
  cat(code_chunk[i])
}
```
=======
cat("\n\n::: {.notes}\n\n")
cat(text_data[8], sep="\n")
cat("\n\n:::\n\n")
```

## 06-02 slide 2

```{r}

cat(text_data[11:23], sep="\n")
```

```{r}
#| output: asis

cat("\n\n::: {.notes}\n\n")
cat(text_data[26], sep="\n")
cat("\n\n:::\n\n")
```

## 06-02 slide 3

```{r}

cat(text_data[29:40], sep="\n")
```

```{r}
#| output: asis

cat("\n\n::: {.notes}\n\n")
cat(text_data[43], sep="\n")
cat("\n\n:::\n\n")
```

## 06-02 slide 4

```{r}

cat(text_data[46:54], sep="\n")
```

```{r}
#| output: asis

cat("\n\n::: {.notes}\n\n")
cat(text_data[57], sep="\n")
cat("\n\n:::\n\n")
```

## 06-02 slide 5

```{r}

cat(text_data[46:80], sep="\n")
```

```{r}
#| output: asis

cat("\n\n::: {.notes}\n\n")
cat(text_data[83], sep="\n")
cat("\n\n:::\n\n")
```

## 06-02 slide 6

```{r}

cat(text_data[86:93], sep="\n")
```

```{r}
#| output: asis

cat("\n\n::: {.notes}\n\n")
cat(text_data[96], sep="\n")
cat("\n\n:::\n\n")
```


## 06-02 slide 7

```{r}

cat(text_data[99:108], sep="\n")
```

```{r}
#| output: asis

cat("\n\n::: {.notes}\n\n")
cat(text_data[111:116], sep="\n")
cat("\n\n:::\n\n")
```
